|

### 2020-12-07 article/article.Rmd
add edits to article, clean up repo

---
title:  |
        | The Effect of Legislature Size on Public Spending
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/distributive-politics-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the law of $1/n$ and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find only mixed evidence for the effect of legislature size on government budgets. Yet the aggregate results mask important differences. While earlier, non-causal studies provide moderate support for the law of $1/n$, recent papers using causal inference methods consistently find a negative relation between seats and spending, suggesting that previous results might have been driven by spurious correlations or omitted variable bias. The available evidence also indicates that XXXXXXXXXXXXXXXX"
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Aggregate results show that legislative size has no significant impact on public
spending. However, the meta-regressions suggest that our study sample has high
levels of heterogeneity, and effects differ substantially according to study
specifications. A reliable predictor of effect size is the research method
employed in the original paper. While papers that employ conventional linear
regressions often find a positive relation between legislature size and
government budgets, recent studies using new methods of causal inference,
specially those with regression discontinuity designs, provide consistent
evidence for a negative relation. This goes in contrast with the original theory
and supports the existence of a "reverse law of $1/n$". Additionally, we find
that non-majoritarian voting systems decrease government spending as a
percentage of GDP, which is also evidence against the "law of $1/n$". In
summary, we do not find support for the theory as originally conceived, and
causal inference models suggest that larger legislatures are associated with
lower public spending.

@@

\begin{longtable}{lccc}\caption{Descriptive Statistics of Moderators\label{tab:descriptive}}\\
\hline  
& Main Coefficients & Other Coefficients &    All Coefficients   \\ 
&       N=42        &       N=100        &    N=142    \\
\hline
\hline     
\endfirsthead 
\multicolumn{4}{l}{\tablename\ \thetable{} \textit{-- continued from previous page}}\\ 
\hline
& Main Coefficients & Other Coefficients &    All Coefficients   \\ 
&       N=42        &       N=100        &    N=142    \\ 
\hline
\hline  
\endhead   
\hline
\multicolumn{4}{l}{\textit{continued on next page}} \\ 
\endfoot   
\multicolumn{4}{l}{}  \\ 
\endlastfoot 
Independent Variables: &                   &                    &            \\ 
$\qquad$K &    12 (28.6\%)     &     33 (33.0\%)     & 45 (31.7\%) \\ 
$\qquad$N &    24 (57.1\%)     &     55 (55.0\%)     & 79 (55.6\%) \\ 
$\qquad$logN &     6 (14.3\%)     &     12 (12.0\%)     & 18 (12.7\%) \\ 
Year &    2009 (6.56)    &    2008 (5.82)     & 2008 (6.04)\\ 
Published work: &                   &                    &            \\ 
$\qquad$Yes &    36 (85.7\%)     &     93 (93.0\%)     & 129 (90.8\%)\\ 
$\qquad$No &     6 (14.3\%)     &     7 (7.00\%)      & 13 (9.15\%) \\ 
Electoral system: &                   &                    &            \\ 
$\qquad$Majoritarian &    22 (52.4\%)     &     46 (46.0\%)     & 68 (47.9\%) \\ 
$\qquad$Non-Majoritarian &    20 (47.6\%)     &     54 (54.0\%)     & 74 (52.1\%) \\ 
Estimation method: &                   &                    &            \\ 
$\qquad$OLS &    11 (26.2\%)     &     36 (36.0\%)     & 47 (33.1\%) \\ 
$\qquad$PANEL &    23 (54.8\%)     &     51 (51.0\%)     & 74 (52.1\%) \\ 
$\qquad$IV &     5 (11.9\%)     &     7 (7.00\%)      & 12 (8.45\%) \\ 
$\qquad$RDD &     3 (7.14\%)     &     6 (6.00\%)      &  9 (6.34\%)  \\

@@

The first two models show the results for public expenditure per capita. No
variable reaches conventional levels of statistical significance in the
restricted sample. In the extended sample, we find that models that use $n$ as
an independent variable tend to detect significantly smaller effects when
compared to $k$. This suggests that an additional member in the lower house has
a smaller impact on public spending than a member in the upper house. Moreover,
the results for the extended sample point out that recent studies find larger
effects than older ones.

@@

Two variables are statistically significant in the last set of
meta-regressions, which include public expenditures as a percentage
of GDP as the dependent variable. Both in our restricted and in our extended
samples, we see that newer studies have smaller coefficients than early ones,

studies with $n$ as an
independent variable have lower coefficients than those that analyse $k$, what
is consistent with our previous models. Both models also show that unpublished
papers tend to have higher coefficients than published papers. Regarding the
electoral systems, see that passing from majoritarian to non-majoritarian
decrease overall levels of public spending. Finally, we find that models
estimated with panel data have larger values than those modelled with OLS.

Overall, our results suggest that study coefficients are highly sensitive to
research design choices. The same statistical methods or study samples may
produce different outcomes depending on the response variables scholars decide
to analyse. Moreover, we find evidence that results vary considerably if the
study employs different measures of legislature size. The impact of factors such
as the electoral system or year of publication also appear to be conditional on
the selected model.

@@

In this article, we assess the empirical validity of the "law of $1/n$". Based
on a sample of `r length(unique(dat$id))` recent publications on the topic, our
meta-analyses show that there is little evidence that an increase in the number
of legislators has a significant effect on public expenditures. If such effect
exists, it is likely driven by an increase in $k$, the size of the upper
legislature, as suggested by several studies in the literature
\citep[e.g.,][]{baqir2002districting, bradbury2001legislative,
bradbury2003local, chen_malhotra_2007, gilligan2001fiscal, primo2006stop}. We
find no robust evidence for the "reverse law of $1/n$", which posits that larger
legislatures lead to lower government spending. One possible explanation is that
the two logics are balancing each other, thus leading us to a null impact of
legislature size on public expenditure.

The meta-regressions indicate that study characteristics have a large influence
on reported results. It is unclear whether factors such as the electoral system
or the level of data aggregation, both believed to moderate the relationship
between legislature size and public expenditure \citep{primo2008distributive,
baqir2002districting, bradbury2003local}, have a substantial impact on the
estimates. We find conflicting evidence for the former and no support for the
latter. Moreover, different statistical techniques produce distinctive results,
with better identified methods (IV and RDD), producing zero and negative
effects more frequently.

One may ask why we find no clear-cut evidence in favour or against the theory.
Although the literature renders simple and testable empirical regularities in both of its
forms, the "law of $1/n$" and the reverse "law of $1/n$", the
empirical assessment is far from uniform in its findings. One reason may be the
fact that the academic reward in social sciences for accumulating evidence is
small, and most papers add features to their theory and further tests while
accessing the "law of $1/n$". For instance, in our sample, papers added to the
theory supermajority rules [@lee2015supermajority; @lee2016supermajority],
redistricting [@baqir2002districting; @lee2018court], party ideology
[@bjedov2014impact], coalition sizes [@baskaran2013coalition], term limits
[@erler2007termlimits], bicameralism [@ricciuti2004legislature], the interplay
between upper and lower houses [@chen_malhotra_2007], among others. This has
the benefit of enriching the original theory with useful details, but has the
disadvantage of providing less clear and decisive tests to the "law of $1/n$"
theory. To solve this problem, we need more direct tests of the theory,
preferably using well causally identified techniques, such as regression
discontinuity design. [XXXX DANILO CHECAR]

Our analyses suggest three areas for further research. First, our study sample
did not include articles that evaluate the association between the log($n$) and
public expenditure per capita or between $k$ and log expenditure per capita.
New work on that area might clarify some of the inconsistencies we find
here. Second, despite the inclusion of several moderators in our models,
aggregate results still show considerable heterogeneity. That is, much of the
disparities among studies are yet to be explained. Domestic factors such as
party dynamics or gerrymandering \citep{lee2015supermajority,
mukherjee2003politicalparties, gilligan2006public} may prove useful in this
regard. Finally, we highlight the need for more evidence in favour or against
the "law of $1/n$" to inform general public policies. The available empirical
evidence points out that contextual variables may either amplify or reduce the
impact of larger legislatures on government budgets. They should be taken into
account if policy-makers want to reach an optimal balance between sound fiscal
policy and the demands for increased political representation.

>>>>

---
title:  |
        | The Effect of Legislature Size on Public Spending
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/distributive-politics-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find no robust evidence for the effect of legislature size on government budgets. Yet the aggregate results show significant heterogeneity. While earlier studies provide moderate support for the \"law of $1/n$\", papers using modern causal inference methods consistently find a negative relationship between seats and spending. The available evidence also indicates that proportional representation and mixed voting systems are no more likely to overspend than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Aggregate results show that legislative size has no significant impact on public
spending. However, the meta-regressions suggest that our study sample has high
levels of heterogeneity, and effects differ substantially according to study
specifications. A reliable predictor of effect size is the research method
employed in the original paper. While papers that employ conventional linear
regressions often find a positive relation between legislature size and
government budgets, recent studies using new methods of causal inference,
specially those with regression discontinuity designs, provide consistent
evidence for a negative relation. This goes in contrast with the original theory
and supports the existence of a "reverse law of $1/n$". Additionally, we find
that non-majoritarian voting systems decrease government spending as a
percentage of GDP, which is also evidence against the theoretical expectations
of the "law of $1/n$". In summary, we do not find support for the theory as
originally conceived, and causal inference models suggest that larger
legislatures are associated with lower public spending.

@@

\footnotesize
\begin{longtable}{lccc}\caption{Descriptive Statistics of Moderators\label{tab:descriptive}}\\
\hline  
& Main Coefficients & Other Coefficients &    All Coefficients   \\ 
&       N=42        &       N=100        &    N=142    \\
\hline
\hline     
\endfirsthead 
\multicolumn{4}{l}{\tablename\ \thetable{} \textit{-- continued from previous page}}\\ 
\hline
& Main Coefficients & Other Coefficients &    All Coefficients   \\ 
&       N=42        &       N=100        &    N=142    \\ 
\hline
\hline  
\endhead   
\hline
\multicolumn{4}{l}{\textit{continued on next page}} \\ 
\endfoot   
\multicolumn{4}{l}{}  \\ 
\endlastfoot 
Independent Variables: &                   &                    &            \\ 
$\qquad$K &    12 (28.6\%)     &     33 (33.0\%)     & 45 (31.7\%) \\ 
$\qquad$N &    24 (57.1\%)     &     55 (55.0\%)     & 79 (55.6\%) \\ 
$\qquad$logN &     6 (14.3\%)     &     12 (12.0\%)     & 18 (12.7\%) \\ 
Year &    2009 (6.56)    &    2008 (5.82)     & 2008 (6.04)\\ 
Published work: &                   &                    &            \\ 
$\qquad$Yes &    36 (85.7\%)     &     93 (93.0\%)     & 129 (90.8\%)\\ 
$\qquad$No &     6 (14.3\%)     &     7 (7.00\%)      & 13 (9.15\%) \\ 
Electoral system: &                   &                    &            \\ 
$\qquad$Majoritarian &    22 (52.4\%)     &     46 (46.0\%)     & 68 (47.9\%) \\ 
$\qquad$Non-Majoritarian &    20 (47.6\%)     &     54 (54.0\%)     & 74 (52.1\%) \\ 
Estimation method: &                   &                    &            \\ 
$\qquad$OLS &    11 (26.2\%)     &     36 (36.0\%)     & 47 (33.1\%) \\ 
$\qquad$PANEL &    23 (54.8\%)     &     51 (51.0\%)     & 74 (52.1\%) \\ 
$\qquad$IV &     5 (11.9\%)     &     7 (7.00\%)      & 12 (8.45\%) \\ 
$\qquad$RDD &     3 (7.14\%)     &     6 (6.00\%)      &  9 (6.34\%)  \\

@@

The first two models show the results for public expenditure per capita. No
variable reaches conventional levels of statistical significance in the
restricted sample. In the extended sample, we find that models that use $n$ as
an independent variable have lower effects when
compared to $k$. This suggests that an additional member in the lower house has
a smaller impact on public spending than a member in the upper house. Moreover,
the results for the extended sample point out that recent studies find larger
effects than older ones.

@@

Two variables are statistically significant in the last set of meta-regressions,
which include public expenditures as a percentage of GDP as the dependent
variable. Both in our restricted and in our extended samples, recent studies
have smaller coefficients than early papers, which stands in contrast with
our previous models. Non-majoritarian voting systems are also
associated with lower levels of public spending, what is against the theoretical
expectations of the "law of $1/n$". The results, however, do not replicate in
the first set of estimations.

Overall, our results suggest that study coefficients are highly sensitive to
research design choices. The same study samples may produce different outcomes
depending on the response variables scholars decide to analyse. Moreover, we
find evidence that results vary considerably if the study employs different
measures of legislature size or use causal research methods. The impact of
factors such as the electoral system or year of publication also appear to be
conditional on the selected model.

@@

In this article, we assess the empirical validity of the "law of $1/n$". Based
on a sample of `r length(unique(dat$id))` publications on the topic, our
meta-analyses show that there is no strong evidence that an increase in the
number of legislators has a significant effect on public expenditures. If such
effect exists, it is likely driven by an increase in $k$, the size of the upper
legislature, as suggested by several studies in the literature
[@baqir2002districting; @bradbury2001legislative; @bradbury2003local;
@chen_malhotra_2007; @gilligan2001fiscal; @primo2006stop]. Instead, we find
better evidence for the "reverse law of $1/n$", which posits that larger
legislatures lead to lower government spending. This is mainly because studies
using regression discontinuity designs, a method that has robust internal
validity, consistently indicate a negative relationship between lower house size
and the logarithm of expenditure per capita [@hohmann2017effect;
@lewis2019legislature; @petterssonlidbom2012size].

The meta-regressions indicate that study characteristics have a considerable
influence on reported results. Factors such as the electoral system affects the
relationship between legislature size and public expenditure, but the results
are not replicable in all estimations. Publication year also produces
conflicting findings in our models. Nevertheless, the meta-regressions confirm
that RDDs produce negative effects more frequently than OLS regressions.

Why is there no clear-cut evidence in favour or against the "law of $1/n$"? A
plausible reason may be that there are few incentives for the pure accumulation
of knowledge in the social sciences, at least when compared to the benefits
scholars may accrue when they challenge or add features to existing theories
[@geddes2003paradigms]. This leads to a reduced number of replication studies in
the field, although we have seen some positive changes in this respect, such as
EGAP's _Metaketa Initiative_.\footnote{See
\url{https://egap.org/our-work/the-metaketa-initiative} for further
information.} For instance, in our sample, papers added supermajority rules
[@lee2015supermajority; @lee2016supermajority], redistricting
[@baqir2002districting; @lee2018court], party ideology [@bjedov2014impact],
coalition sizes [@baskaran2013coalition], term limits [@erler2007termlimits],
bicameralism [@ricciuti2004legislature], and the interplay between upper and
lower houses [@chen_malhotra_2007] to the main theory, but a comprehensive
procedural replication of the findings using different samples is yet to be
written. The addition of new features has the benefit of enriching the original
theory with useful details, although it has the disadvantage of not providing a
conclusive test to the "law of $1/n$".

Our analyses suggest three areas for further research. First, our study
sample did not include articles that evaluate the association between the
log($n$) and public expenditure per capita or between $k$ and log expenditure
per capita. New work on that area might clarify some of the inconsistencies we
find here. Second, despite the inclusion of several moderators in our models,
aggregate results still show considerable heterogeneity. Domestic factors such
as party dynamics or gerrymandering [@lee2015supermajority;
@mukherjee2003politicalparties; @gilligan2006public] may prove useful in this
regard. Finally, we highlight the need for more causal inference studies in the
literature. Whenever possible, authors should leverage natural and
quasi-experiments to assess whether the current results hold when tested with
such research designs. These suggestions may help scholars and policy-makers to
reach an optimal balance between sound fiscal policy and the demands for
increased political representation.
